Modeling browser-based distributed evolutionary computation systems

### 2015-02-08 modeling-volunteer-computing-gecco.tex
First look to the paper and a possible reference to use added
Some comments have been added and also some typos fixed.

\begin{abstract}
From the era of big science we are back to the "do it yourself", where
you do not have any money to buy clusters or subscribe to grids but
still have algorithms that crave many computing nodes and need them to
measure scalability. Fortunately, this coincides with the era of big
data, cloud computing, and browsers that include JavaScript virtual
machines. Those are the reasons why this paper will focus on two
different aspects of volunteer or freeriding computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to a environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a very simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

Most theory in parallel computing has been devoted to predict and
optimize the performance in systems where the number of nodes, their
connections, and the time every one is devoting to the computation is
known in advance. However, even if Big Science is not really over and
it is slated to come back, the era of
Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
offers a vast amount of computational resources to attract, if only
you know how. But there is a challenge: knowing, or at least having a
ballpark, of how your algorithm is going to perform in this uncertain
environment, where none of the factors is known: neither the number of
nodes, through how they are connected, to how long are they going to
be focused on doing what you want them to. That is why some effort is
needed to first understand the dynamics of the decision to participate
in an experiment that requires you to click on a link and then stay
for a while looking at the screen (or just leave it there running).

Besides, since Amazon started selling EC2 several years ago, reliable
and scalable computing resources are also available for a low price
and on demand. Recently, Google has also refurbished its offering
lowering their prices. This means that the conjunction of free or
low-cost cloud computing engines, volunteer computing systems and and
untapped capability of desktop systems can be used for creating
massive, or at least potentially massive, distributed computing
experiments. These experiments can be easily created using open-source
repository sites like GitHub and deployed automatically to Platform as
a Service products such as Heroku or OpenShift.

In general, any volunteer computing experiment will have to be made
``in the open''. The fact that somebody is giving you, basically for
free, computing resources, means that the scientist using them has to
give back. The baseline is releasing the source code used: all
volunteer computing platforms, from SETI@home on, have done it. The
opposite is probably the reason why many companies like PopularPower \cite{buyya2001compute}
have folded or others like CrowdProcess have decided to turn their
product to in-house computation: there must be a mutual relationship
of trust among the scientist and the person that is running his/her
code in the browser. As has been mentioned in the early stages of what
was then called {\em desktop grid computing}, \cite{gc:bausch}, in
fact CPU cycle selling might not make economic sense since there are
not so much demand for it and supply is quite high. However, while
potential supply is in fact huge, {\em actual} supply will depend on
the person holding it willing to actually allocate it to a particular
company selling it or a particular experiment needing it, not to
mention the fact that the experiment {\em actually} has to draw the
attention of the supplier. That is why trust is essential, and using
free software might not be enough: Openness
has to progress from open source code to open science: releasing all
data obtained in the experiment in a repository such as GitHub,
mentioned above, and even allowing real-time access to experiment data
to users.

@@

These are best practices that have been followed in the experiments in
this paper, that first presents the first versions of a platform for
distributed evolutionary algorithms using the browser and a free (as
in free beer) backend, and second, shows the result from an
statistical point of view in order to put the basis for a model of the
metacomputer obtained by joining all volunteers and the free backend
used for the experiment. This is not an exhaustive or complete
exploration of the possibilities of this kind of computation, but it
is enough to present the free software used to perform the experiments
and will allow us to describe, in general, the behavior of the users
as well as the performance achieved on the experiments done so far,
which should show some advantage over doing the same kind of
experiments locally using available resources.

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelization. We
will proceed to describe the experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'' and
\begin{quote}
Java can bring some important advantages... solves architectural
heterogeneity [...] easy to install and [...] built-in security
mechanisms
\end{quote}
The same paper by Soares et al. describes JET, a system that supports
the execution of parallel applications over the Internet. In fact the
same paper describes some other systems working at the same time,
except they cannot be used for science since they do not have a
comprehensive statistics support like JET.

However, Java (and all the other technologies) is not universal in the
sense than an extra component, namely, the Java virtual machine, has
to be installed in the browser. JavaScript
\cite{flanagan2006javascript} was introduced in 1997 as a
browser-embedded language and has, since them, become a set of standards
\cite{ECMA-262} for the language, its components and future versions
that have been widely adopted by the industry and also by scientists,
who used them for creating a non-distributed evolutionary algorithm on
the browser as early as 1998 \cite{jj-ppsn98}.

The potential for volunteer computing using browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential for
mischief of users with code in their hands
\cite{sarmenta-sabotagetolerance} however, these early efforts by
Sarmenta once again used Java and not JavaScript, making this effort
less universal. JavaScript can be used either for unwitting
\cite{unwitting-ec} or volunteer
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and has been used ever since by several
authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
used the client's GPU \cite{duda2013gpu}. Many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
platform avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

There have been relatively few efforts to analyze what is the
performance that can be obtained from these volunteer computing
effort. There was some effort initially to dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the evolutionary algorithm that can easily be
paralelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of
performance of the evolutionary algorithm itself
\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
volunteer computing, number of users and the time spent in the
computation in browser-based volunteer computing experiments, have only been studied in a limited way in
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}.

As far as we know, this is the only experiment that uses computational
resources that are as dissimilar as smartphones and powerful laptops
or desktop computers in a research center. The methodology used to
gather resources and the algorithms used will be described next.

@@

\begin{figure*}[h!tb]
  \centering
  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
  \label{fig:sshot}
  \caption{Screenshot of the talk that includes at the graph
    that shows the progression of the evolutionary
    algorithm at the top right corner}
\end{figure*}

In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences\footnote{Names withheld for the double blind review} and
it has the appearance shown in Figure \ref{fig:sshot}. During the
conference, it was revealed to the users that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. As it can be seen in the
image, the upper right corner shows the progress of the evolution. It
stops when the solution has been reached.

Above the experiment, from the point of view of the user, has been
described. What it is actually running is an island model
\cite{muhlenbein1991parallel} in every browser that uses the server as
a {\em shuttle} to transfer individuals from one browser to another in
what can be eventually a fully connected topology. This deals with
several problems. The connection is stateless: islands connect to the
server to send and receive a single individual but there is no task
balancing: all islands start clicking on an URL and finish when they
browse off to the next page. Fault tolerance is implicit, in the sense
that no island has information that cannot be, or generated, somewhere
else, although if the server fails the experiment results might be
lost. All operation is asynchronous, with islands entering and leaving
the experiment all of their own.

@@

This fitness is implemented as part of a (name hidden) evolutionary
algorithm library written in node.js, a server-based implementation of
JavaScript. However, since it is going to be run on the browser, a
procedure called ``browserification'' must be applied to it so that it
can run, without modifications, in any browser. The main difference
between JavaScript on the browser and node is the way they load the
external libraries. The algorithm used a population of 256
individuals, random initialization, a canonical evolutionary algorithm
with an elite of 2 individuals and crossover and then mutation applied
to all the population pool, that was generated using 2-individual
tournament selection. Once again, the particular evolutionary
algorithm used is not so important, since we were rather focused on
the user behavior, although it is quite clear that the selective
pressure of the algorithm will have an impact in its performance and
will play for or against its asynchrony in many different and complex
ways \cite{jj:2008:PPSN}.

The initial code was only modified to represent the fitness in a chart
(that uses the {\tt chart.js} JavaScript library) so that the user has
a visual feedback of what is going on. The {\em migration} part is
also added: after every 100 generations, a single individual is sent
to the server and another request is made to the server of a single
and random individual. This makes the procedure less greedy, and also
decreases the probability of getting the same individual that was sent
some generations back. These two operations are made in {\em fire and
  forget} style without any error checking, so that if the server is
down, or indeed if the algorithm is run locally, it is not interrupted
and continues operation until the solution is found, a string with all
ones.

@@

In fact, the program includes more logic to serve the static pages
that contain the talk in which it is included and the JavaScript/CSS
code also, but it all amounts to a small amount of code. This code can
be deployed, in principle, in any server running node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a freeriding way, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014,jj:idc:lowcost}.

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in a GitHub
repository (address hidden for double blind review). These results
will be analyzed in the next section.

@@

The litmus test that all experiments should pass, that is, that they
work and found the solution, passed. In fact, the first version of the
algorithm (until January 2015) used too much memory on the browser,
which crashed after a while running. A total of 7 experiments were
done with this version. After January, a new version of the
evolutionary algorithm with non-cached fitness was used which did not
crash the browser and was thus allowed to run as much time as
needed. That was the version used in a lightning talk at the beginning
of February. A total of 10 experiments were made with this second
version. In the first version, finding the solution took a few
minutes. The second needed around 20 minutes. In the logs used, it was
not checked that the algorithm had found the solution, but in most
cases it had. More systematic experimentation would have to wait until
the next version, since in this first one we were looking for some
kind of systematicity in the behavior of users and the experimental
apparatus is the minimum possible to have it running.
%
\begin{figure*}[htb]
        \centering
        \subfigure[First experiment]{
                \includegraphics[width=0.45\linewidth]{data/diffs-ips-1.png}
                \label{fig:e1_p}
        }
        \subfigure[Second experiment]{
                \includegraphics[width=0.45\linewidth]{data/diffs-ips-2.png}
                \label{fig:e2_p}
        }
        \caption{Histogram of number of different IPs in an experiment}\label{fig:histo:ips}
\end{figure*}

The fact that it run correctly during a massive lightning talk means that this setup is a
valid, no-cost, platform for massive evolutionary algorithms. There
were several ways of introducing it: during a talk, telling the
audience to visit the web that was posted in the first slide, and
after the talk, inserting the URL in a tweet with a brief explanation
and several tweets more indicating what it was about and what it
did. If any question arose, it was answered as fast as possible.

@@

There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}. But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public. This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.

>>>>

\begin{abstract}
From the era of big science we are back to the ``do it yourself'', where
you do not have any money to buy clusters or subscribe to grids but
still have algorithms that crave many computing nodes and need them to
measure scalability. Fortunately, this coincides with the era of big
data, cloud computing, and browsers that include JavaScript virtual
machines. Those are the reasons why this paper will focus on two
different aspects of volunteer or freeriding computing: first, the
pragmatic: where to find those resources, which ones can be used, what
kind of support you have to give them; and then, the theoretical: how
evolutionary algorithms can be adapted to a environment in which nodes
come and go, have different computing capabilities and operate in
complete asynchrony of each other. We will examine the setup needed to
create a very simple distributed evolutionary algorithm using
JavaScript and then find a model of how users react to it by
collecting data from several experiments featuring different classical
benchmark functions. 
\end{abstract}

@@

Most theory in parallel computing has been devoted to predict and
optimize the performance in systems where the number of nodes, their
connections, and the time every one is dedicating to the computation is
%FERGU: cambiado el devoting que se repite, por dedicating 
known in advance. However, even if Big Science is not really over and
it is slated to come back, the era of
Citizen Science already started a few years ago (with SETI@home \cite{david-seti:home} and then BOINC \cite{boinc_grid04}) and it
offers a vast amount of computational resources to attract, if only
you know how. But there is a challenge: knowing, or at least having a
ballpark, of how your algorithm is going to perform in this uncertain
environment, where none of the factors is known: neither the number of
nodes, through how they are connected, to how long are they going to
be focused on doing what you want them to. That is why some effort is
needed to first understand the dynamics of the decision to participate
in an experiment that requires you to click on a link and then stay
for a while looking at the screen (or just leave it there running).

Besides, since Amazon started selling EC2 several years ago, reliable
and scalable computing resources are also available for a low price
and on demand. Recently, Google has also refurbished its offering
lowering their prices. This means that the conjunction of free or
low-cost cloud computing engines, volunteer computing systems and and
untapped capability of desktop systems can be used for creating
massive, or at least potentially massive, distributed computing
experiments. These experiments can be easily created using open-source
repository sites like GitHub\footnote{\url{http://www.github.com}} and deployed automatically to Platform as
a Service (PaaS) products such as Heroku\footnote{\url{https://www.heroku.com/}} or OpenShift\footnote{\url{https://www.openshift.com/}}. 
%FERGU: luego usas el acrónimo PaaS

In general, any volunteer computing experiment will have to be made
``in the open''. The fact that somebody is giving you, basically for
free, computing resources, means that the scientist using them has to
give back. The baseline is releasing the source code used: all
volunteer computing platforms, from SETI@home on, have done it. The
opposite is probably the reason why many companies like PopularPower \cite{buyya2001compute}
have folded or others like CrowdProcess have decided to turn their
product to in-house computation: there must be a mutual relationship
of trust among the scientist and the person that is running his/her
code in the browser. As it has been mentioned in the early stages of what
was then called {\em desktop grid computing} \cite{gc:bausch}, in
fact CPU cycle selling might not make economic sense since there are
not so much demand for it and supply is quite high. However, while
potential supply is in fact huge, {\em actual} supply will depend on
the person holding it willing to actually allocate it to a particular
company selling it or a particular experiment needing it, not to
mention the fact that the experiment {\em actually} has to draw the
attention of the supplier. That is why trust is essential, and using
free software might not be enough: Openness
has to progress from open source code to open science: releasing all
data obtained in the experiment in a repository such as GitHub,
mentioned above, and even allowing real-time access to experiment data
to users.

% Había metido en el .bib el artículo de detectar vulnerabilidades 
% en EAs distribuidos de Juanlu, pero creo que al final no tiene sentido 
% hablar en este paper de él (Muszynski14vulnerabilities)

@@

These are best practices that have been followed in the experiments in
this paper, that first presents the first versions of a platform for
distributed evolutionary algorithms (EAs) using the browser and a free (as
in free beer) backend, and second, shows the result from an
statistical point of view in order to put the basis for a model of the
metacomputer obtained by joining all volunteers and the free backend
used for the experiment. This is not an exhaustive or complete
exploration of the possibilities of this kind of computation, but it
is enough to present the free software used to perform the experiments
and will allow us to describe, in general, the behavior of the users
as well as the performance achieved on the experiments done so far,
which should show some advantage over doing the same kind of
experiments locally using available resources.

The rest of the paper is organized as follows: next we will present
the state of the art in volunteer computing and its modelization. We 
%FERGU: aunque hay artículos con el término "modelization", no sé si el correcto será "modelling"
will proceed to describe the experimental setup in Section
\ref{sec:exp}, some preliminary results in Section \ref{sec:res} and
finally we will present our conclusions in Section \ref{sec:conc}
along with future lines of work.

@@

Using the web as a resource for distributed, or plainly user based,
evolutionary computation has a long history since JavaScript was
introduced as a browser-based language in 1997 and even before, when
other procedures, including Flash animations, VBScript, ActiveX or Java applets were
used. Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms.
%\begin{quote}
%Java can bring some important advantages... solves architectural
%heterogeneity [...] easy to install and [...] built-in security
%mechanisms
%\end{quote} FERGU: quito esta cita y lo pongo sin el quote
In that work, Soares et al. describe JET, a system that supports
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science, 
unlike the other systems compared in that paper.

However, Java (and all the other technologies)
is not universal in the
sense than an extra component, namely, the Java virtual machine, has
to be installed in the browser. JavaScript
\cite{flanagan2006javascript} was introduced in 1997 as a
browser-embedded language and has, since them, become a set of standards
\cite{ECMA-262} for the language, its components and future versions
that have been widely adopted by the industry and also by scientists,
who used them for creating a non-distributed EA on
the browser as early as 1998 \cite{jj-ppsn98}.

The potential for volunteer computing using browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential for
mischief of users with code in their hands
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making this effort
less universal. JavaScript can be used either for unwitting
\cite{unwitting-ec} or volunteer
\cite{langdon:2005:metas,gecco07:workshop:dcor} distributed
evolutionary computation and it has been used ever since by several
authors, including more recent efforts \cite{Desell:2008:AHG:1389095.1389273,duda2013distributed,DBLP:journals/corr/abs-0801-1210} that even
used the client's GPU \cite{duda2013gpu}. Many other researchers have
used Java \cite{chong:1999:jDGPi} and others have gone away from the
server-based paradigm to embrace peer to peer systems
\cite{jin2006constructing,10.1109/ICICSE.2008.99}. These computing
platform avoid single points of failure (the server) but, since they
need a certain amount of infrastructure installed to start, the
threshold to join them is much lower.

There have been relatively few efforts to analyze what is the
performance that can be obtained from these volunteer computing
effort. There was some attempt initially to dodge the issue by making
the algorithm adaptive to the kind of resources allotted to it
\cite{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can easily be
parallelized via population splitting or farming out evaluations to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer algorithms
\cite{gonzalez2010characterizing} which can, of course, be studied in
a more general distributed computing context
\cite{nogueras2015studying} or including it in a more general study of
performance of the EA itself %Con tu permiso lo paso a acrónimos para que no sea tan largo el párrafo. FERGU
\cite{DBLP:journals/gpem/LaredoBGVAGF14}. But the raw material of
volunteer computing, number of users and the time spent in the
computation in browser-based volunteer computing experiments, have only been studied in a limited way in
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and Gamma distribution
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}.

As far as we know, this paper presents the only experiment that uses computational
resources that are as dissimilar as smartphones and powerful laptops
or desktop computers in a research center. The methodology used to
gather resources and the algorithms used will be described next.

%FERGU: volver a indicar que además se dan resultados de uso, para que quede clara la contribución?

@@

\begin{figure*}[h!tb]
  \centering
  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
  \label{fig:sshot}
  \caption{Screenshot of the talk that includes at the graph
    that shows the progression of the evolutionary
    algorithm at the top right corner}
\end{figure*}
%FERGU: No puedes poner el gráfico un poco más grande? (y de paso 
% recortar mejor la imagen para que no se vea el texto de la derecha)
In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences\footnote{Names withheld for the double blind review} and
it has the appearance shown in Figure \ref{fig:sshot}. During the
conference, it was revealed to the users that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. As it can be seen in the
image, the upper right corner shows the progress of the evolution. It
stops when the solution has been reached.

Above the experiment, from the point of view of the user, has been
described. What it is actually running is an island model
\cite{muhlenbein1991parallel} in every browser that uses the server as
a {\em shuttle} to transfer individuals from one browser to another in
what can be eventually a fully connected topology. 
%FERGU: explica esto: los navegadores se conectan entre sí o sólo a tu servidor?
This deals with
several problems. The connection is stateless: islands connect to the
server to send and receive a single individual but there is no task
balancing: all islands start clicking on an URL and finish when they
browse off to the next page. Fault tolerance is implicit, in the sense
that no island has information that cannot be, or generated, somewhere
else, although if the server fails the experiment results might be
lost. All operation is asynchronous, with islands entering and leaving
the experiment all of their own.

@@

This fitness is implemented as part of a (name hidden) evolutionary
%FERGU: creo que si usas un framework libre puedes dejarlo visible
algorithm library written in node.js, a server-based implementation of
JavaScript. However, since it is going to be run on the browser, a
procedure called ``browserification'' must be applied to it so that it
can run, without modifications, in any browser. The main difference
between JavaScript on the browser and node.js is the way they load the
%FERGU: puesto .js para distinguir la palabra node del nombre de la librería
external libraries. The algorithm used a population of 256
individuals, random initialization, a canonical evolutionary algorithm
with an elite of 2 individuals and crossover and then mutation applied
to all the population pool, that was generated using 2-individual
tournament selection. Once again, the particular evolutionary
algorithm used is not so important, since we were rather focused on
the user behavior, although it is quite clear that the selective
pressure of the algorithm will have an impact in its performance and
will play for or against its asynchrony in many different and complex
ways \cite{jj:2008:PPSN}. %FERGU: este párrafo me lo copio para el de la ASOCO ;)

The initial code was only modified to represent the fitness in a chart
(that uses the {\tt chart.js} JavaScript library) so that the user has
a visual feedback of what is going on. The {\em migration} part is
also added: after every 100 generations, a single individual is sent
to the server and another request is made to the server of a single
and random individual. This makes the procedure less greedy, and also
decreases the probability of getting the same individual that was sent
some generations back. These two operations are made in {\em fire and
  forget} style without any error checking, so that if the server is
  %FERGU: Qué es fire and forget? (se puede intuir, pero dejarlo claro)
down, or indeed if the algorithm is run locally, it is not interrupted
and continues operation until the solution is found, a string with all
ones.

@@

In fact, the program includes more logic to serve the static pages
that contain the talk in which it is included and the JavaScript/CSS
code also, but it all amounts to a small amount of code. This code can
be deployed, in principle, in any server running node.js, but for the
purposes of this experiment Heroku, a Platform as a Service with a
free tier, has been chosen. It could be also be deployed, with small
or no modification, to OpenShift or other similar PaaS. Far from being
a freeriding way, using free or very low cost servers can be a
sustainable way of performing massive distributed evolutionary
computation experiments \cite{Merelo2014}

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in a GitHub
repository (address hidden for double blind review).

@@

@@

There are many issues involved in using these resources: from adapting current algorithms so that they match this environment \cite{agajaj} to check which EA configuration works the best in it, or creating a framework that can use it easily \cite{nodeo2014}, or even security issues. But the main challenge is that asking people to contribute resources implies that you are opening your science to society and you have to give something in return: you have to adopt a set of best practices that have come to be known as Open Science, because ``Give, and it shall be given unto you'', you will get as much back from society as you give to it opening your science and explaining it to the public. This, among other things, means that popularity will become directly performance of the metacomputer you create by attracting more users.
